---
title: Testing Equivalence between NIRS Breakpoints and Ventilatory Thresholds
subtitle: Computational Notebook
author:
  - name: Philip Skotzke
    orcid: 0000-0003-4474-4637
    email: s8phskot@uni-saarland.de
    affiliation:
      - name: Department of Movement, Human and Health Sciences, University of Rome "Foro Italico", Rome, Italy
        city: Rome
date: today
format:
  html:
    toc: true
    toc-location: right-body
    toc-depth: 3
    toc-title: Contents
    embed-resources: true
    self-contained-math: true
    code-fold: true
    other-links: 
     - text: Github 
       href: https://github.com/PhilipSkotzke/breakpoints_equivalence
     - text: OSF
       href: https://osf.io/u8k2e/
    theme: lumen
    number-sections: false
    #number-depth: 2
bibliography: data/references/references.bib
csl: data/references/apa.csl
tbl-cap-location: top
---

```{r}
#| label: SetUp
#| warning: false

# Set up for the computational notebook
# -> load packages
# -> set the base ggplot2 theme
# -> define custom colors
# -> load extra fonts
# -> save package version with {renv}

library(dplyr)      # efficient data manipulation
library(extrafont)  # fonts for publication plots
library(extrafontdb)# database for extra fonts
library(ggdist)     # plot distributions
library(ggplot2)    # create plots
library(ggtext)     # more flexible plot texts
library(gt)         # gt tables
library(gtsummary)  # summary tables
library(kableExtra) # create html tables
library(MetBrewer)  # better color palettes
library(patchwork)  # combine plots
library(readxl)     # read MS Excel files
library(renv)       # save the project environment
library(stringr)    # manipulate character strings
library(tidyr)      # data management
library(TOSTER)     # perform two onesided tests


#### set up the base theme ####
theme_set(
    theme_bw(base_size = 14) %+replace%
        theme(
            plot.title = ggtext::element_markdown(
                colour = "grey10", size = rel(1.2), hjust = 0, 
                lineheight = 1.2, margin = margin(b = 7)),
            plot.title.position = "plot",
            plot.caption.position = "plot",
            panel.border = element_rect(colour = "black",
                                        fill = NA, linewidth = 0.7),
            axis.line = element_blank(),
            panel.grid.major = element_blank(),
            panel.grid.minor = element_blank(),
            text = element_text(colour = "grey20"),
            axis.text = element_text(colour = "grey20"),
            axis.title.x = element_text(margin = margin(t = 7)),
            axis.title.y.right = element_text(angle = -90,
                                              margin = margin(l = 4)),
            axis.title = element_text(face = "bold"),
            legend.margin = margin(c(1, 1, 1, 1)),
            legend.position = "top",
            legend.justification = "right",
            legend.direction = "horizontal",
            legend.key = element_blank(),
            legend.key.width = unit(0.7, 'cm'),
            legend.spacing.y = unit(0.2, 'cm'),
            strip.background = element_rect(fill = "grey95",
                                            colour = "grey10",
                                              linewidth = 0.7),
            strip.text = element_text(
                size = rel(1), margin = margin(4, 4, 4, 4)),
            complete = FALSE
        ))


#### define custom colors ####
clrs <- met.brewer("Juarez")


#### load extra fonts ####
# run only if fonts haven't been imported before

# font_import(prompt = T)

# load fonts
loadfonts(quiet = T)

#### create reproducible envvironments with {renv} ####
# run only once, then comment out

# init()

# update renv files (comment out when not needed)

#snapshot()
```

## Preface

This file is the **computational notebook** accompanying the main manuscript and is stored on [Github](https://github.com/PhilipSkotzke/breakpoints_equivalence "Link to GitHub Repository") and accessible on [OSF](https://osf.io/u8k2e/ "Link to OSF project"). It contains the reasoning behind analyses performed, the full details on statistical procedures and the analytic code. Please note, that:

-   Sub-headers in @sec-manuscript refer to figures and tables generated for the main manuscript or supplements. They are stored in their respective folders `/output/results` and `/output/supplementary` and not presented here.

-   The code used to produce the results in this document can be viewed by clicking on the `Code` button to unfold it.

::: {.callout-tip collapse="true"}
## This is a reproducible document

This `.html` file is technically reproducible. Download the repository from OSF or GitHub and run the file `analysis_breakpoints_equivalence.qmd`.
:::

## Hypothesis

To be transparent about our reasoning process, we make the distinction between the *research question*, the *research hypothesis*, the *empirical hypothesis* and the *statistical hypothesis* (for details read [here](https://publish.obsidian.md/methodology-swi-uds/Practise/Analyzing+research+methodology/1.+Find+the+research+question)) and state each separately.

::: callout-note
This is an exploratory analysis [@Ditroilo2025]. We use existing data to test a hypothesis, but no adequate error control can be ensured. Instead, we use our results to estimate the required sample size to reach sufficient statistical power when comparing NIRS breakpoints with ventilatory thresholds. We believe that this will help to inform future research to recruit adequate sample sizes for confirmatory studies comparing NIRS breakpoints with a criterion method.
:::

### Research Question

Can we use NIRS-measured muscle oxygen saturation (SmO~2~) to delimit the three exercise intensity domains in a field based running test?

### Research Hypothesis

1.  BP1 can be used interchangeably with GET.
2.  BP2 can be used interchangeably with RCP.

Note, based on the existing literature no directional difference would be assumed (e.g. it is as likely that BP2 is at a higher intensity than RCP than the other way around).

### Empirical Hypothesis

1.  The heart rate and speed derived from a the VAM-EVAL incremental test in highly trained male triathletes are equivalent between BP1 and GET .
2.  The heart rate and speed derived from a the VAM-EVAL incremental test in highly trained male triathletes are equivalent between BP2 and RCP.

### Statistical Hypothesis

#### BP1 vs GET

$\text{H}_0: \left|\mu_{\text{GET.HR}} - \mu_{\text{BP1.HR}} \right| \geq \Delta_{\text{HR}} ~ \land ~ \left|\mu_{\text{GET.SPEED}} - \mu_{\text{BP1.SPEED}} \right| \geq \Delta_{\text{SPEED}}$

$\text{H}_1: \left|\mu_{\text{GET.HR}} - \mu_{\text{BP1.HR}} \right| < \Delta_{\text{HR}} ~ \land ~ \left|\mu_{\text{GET.SPEED}} - \mu_{\text{BP1.SPEED}} \right| < \Delta_{\text{SPEED}}$

#### BP2 vs RCP

$\text{H}_0: \left|\mu_{\text{RCP.HR}} - \mu_{\text{BP2.HR}} \right| \geq \Delta_{\text{HR}} ~ \land ~ \left|\mu_{\text{RCP.SPEED}} - \mu_{\text{BP2.SPEED}} \right| \geq \Delta_{\text{SPEED}}$

$\text{H}_1: \left|\mu_{\text{RCP.HR}} - \mu_{\text{BP2.HR}} \right| < \Delta_{\text{HR}} ~ \land ~ \left|\mu_{\text{RCP.SPEED}} - \mu_{\text{BP2.SPEED}} \right| < \Delta_{\text{SPEED}}$

Where $\Delta$ denotes the equivalence range.

As indicated by the formulation, the hypothesis can be tested using an equivalence test [@Lakens2017]. Based on the existing literature and practical relevance, the following values will be used as smallest effect size of interest (SESOI):

The test-retest reliability for speed at the GET and RCP is typically in the range of 0.2-0.4 km/h [e.g. @Cerezuela-espejo2018; @Dickhuth1999]. Furthermore, a critical difference of 0.5 km/h between lactate thresholds and MLSS has been used before for criterion validity and is justified by its relevance to athletic performance [@Dörr2010]. **Thus, we will use 0.5 km/h as SESOI.**

For heart rate, the typical error of measurement has been reported to be 3.8% for VT1 and 2.5% for VT2 [@Weston2001], corresponding to 5 and 4 bpm, respectively. In a different investigation, the standard error of measurement was 3-4 bpm [@Yogev2023b]. **Thus, we will use 5 bpm as SESOI.**

Please note, the two hypotheses pairs are formulated as conjunction. In contrast to @Mesquida2025 we use logical connectors instead of set operators. We argue that the intersection-union approach is the right one for our research question. It would be justified to go a step further and use the $\land$ ("and") connector to combine our hypotheses about the first and second threshold because only if we can use SmO~2~ to delineate *all* three intensity domains it would be interchangeable with ventilatory thresholds. However, if NIRS can be used to establish either threshold, it would be practically useful. As a consequence, we don't have to apply $\alpha$ error corrections.

For all analysis we use $\alpha$ \< 0.05 for Type I error control as it is common practice in our field.

## Data Analysis

The data used for this analysis was generated as part of a different project, where the VAM-EVAL was performed as initial test to assess participant fitness. See the codebook (`data/input data/metadata`) for more details.

### Data Preparation

The data has been been pre-processed and only needs to be loaded.

```{r}
#| label: LoadData

# load the data
Study_data <- read_xlsx("data/input data/data_NIRS_BPs_2025.xlsx")
```

### Descriptive Statistics

First, we create tables with descriptive statistics.

```{r}
#| label: DescStat

# the gtsummary::tbl_summary() function is used after selecting the variables for participant characteristics
Study_data %>%
  select(Age, Height, Weight, ATT_left, ATT_right, v_peak, MAS, absVO2max, VO2max) %>%
  mutate(BMI = Weight / (Height/100)^2) %>%
    tbl_summary(
    type = ~ "continuous",
    statistic = all_continuous() ~ "{mean} \u00B1 {sd} [{min} - {max}]",
    missing_text = "missing",
    digits = list(
      c(Age, Height, Weight, ATT_left, ATT_right, v_peak, MAS, VO2max, BMI) ~ 1,
      c(absVO2max) ~ 0
    )
  ) %>%
  as_gt() %>%
  tab_header(title = "Participant Characteristics")

# then the variables relating to the threshold identification are selected and summarized
Study_data %>%
  select(-c(ID, Age, Height, Weight, ATT_left, ATT_right, v_peak, MAS, VO2max, absVO2max)) %>%
  tbl_summary(
    type = ~ "continuous",
    statistic = all_continuous() ~ "{mean} \u00B1 {sd}",
    missing_text = "not identified",
    digits = list(
      c(GET_speed, RCP_speed, BP1_speed, BP2_speed) ~ 1,
      c(GET_time, GET_HR, RCP_time,
        RCP_speed, BP1_time, BP1_speed, BP2_time, BP2_speed) ~ 0
    )
  ) %>%
  as_gt() %>%
  tab_header(title = "Threshold Identification Details")
  
```

### Inferential Statistics

To test the assumptions, the normality of differences is visually checked using qq-plots and statistically tested with the Shapiro-Wilk test for heart rate and speed at the first and second threshold.

#### First Threshold (GET vs. BP1)

##### Assumptions - heart rate

```{r}
#| label: AssumptionsThd1HR

# calculate the difference
diff_HR1 <- Study_data$GET_HR - Study_data$BP1_HR 

# generate the qq plot
qqnorm(diff_HR1)

# perform the Shapiro-Wilk test
shapiro.test(diff_HR1)
```

The assumption of normality of the differences is not violated for heart rate.

##### TOST heart rate first threshold

```{r}
#| label: TOSTThd1HR

# the t_TOST function is used
EQ1_HR <- t_TOST(
  x = Study_data$GET_HR,
  y = Study_data$BP1_HR,
  hypothesis = "EQU",
  paired = T,
  #var.equal = T,
  eqb = 5,
  alpha = 0.05
)

# print the results
EQ1_HR
plot(EQ1_HR, "simple", estimates = "raw")
```

##### Assumptions speed

```{r}
#| label: AssumptionsThd1Speed

# calculate the difference
diff_SP1 <- Study_data$GET_speed - Study_data$BP1_speed 

# create the qq-plot
qqnorm(diff_SP1)

# perform the Shapiro-Wilk test
shapiro.test(diff_SP1)
```

The assumption of normality of the differences is not violated for speed.

##### TOST speed first threshold

```{r}
#| label: TOSTThd1Speed

# perform the TOST
EQ1_SP <- t_TOST(
  x = Study_data$GET_speed,
  y = Study_data$BP1_speed,
  hypothesis = "EQU",
  paired = T,
  #var.equal = T,
  eqb = 0.5,
  alpha = 0.05
)

# print the results
EQ1_SP
plot(EQ1_SP, "simple", estimates = "raw")
```

#### Second Threshold (RCP vs. BP2)

##### Assumptions - heart rate

```{r}
#| label: AssumptionsThd2HR

# the steps are the same as above

diff_HR2 <- Study_data$RCP_HR - Study_data$BP2_HR 

qqnorm(diff_HR2)

shapiro.test(diff_HR2)
```

The assumption of normality of the differences is not violated for heart rate.

##### TOST heart rate second threshold

```{r}
#| label: TOSTThd2HR

EQ2_HR <- t_TOST(
  x = Study_data$RCP_HR,
  y = Study_data$BP2_HR,
  hypothesis = "EQU",
  paired = T,
  #var.equal = T,
  eqb = 5,
  alpha = 0.05
)

EQ2_HR
plot(EQ2_HR, "simple", estimates = "raw")
```

##### Assumptions - speed

```{r}
#| label: AssumptionsThd2Speed

diff_SP2 <- Study_data$RCP_speed - Study_data$BP2_speed

qqnorm(diff_SP2)

shapiro.test(diff_SP2)
```

The assumption of normality of the differences is not violated for speed.

##### TOST speed second threshold

```{r}
#| label: TOSTThd2Speed

EQ2_SP <- t_TOST(
  x = Study_data$RCP_speed,
  y = Study_data$BP2_speed,
  hypothesis = "EQU",
  paired = T,
  #var.equal = T,
  eqb = 0.5,
  alpha = 0.05
)

EQ2_SP
plot(EQ2_SP, "simple", estimates = "raw")
```

### Power Analysis/Sample Size Estimation

Initially we set out to perform a sensitivity analysis to calculate the statistical power we achieved with our sample. However, as the mean differences are outside our pre-defined equivalence range, we cannot calculate our power (it is 0).

What we can do instead is to calculate the sample size needed to achieve 80% power with an $\alpha$ of 0.05, our pre-defined equivalence bounds *if a breakpoint estimation method is used that has a true mean difference of 0 compared to the ventilatory threshold*, and the standard deviation of the difference is reasonably well estimated by the standard deviation of our sample.

```{r}
#| label: SampleSizeEst

# sample size calculation for true zero difference
power_EQ1_HR <- power_t_TOST(
  alpha = 0.05,
  power = 0.80,
  delta = 0,
  sd = EQ1_HR$effsize$SE[[1]]*sqrt(EQ1_HR$smd$ntilde),
  eqb = 5,
  type = "paired"
)

power_EQ2_HR <- power_t_TOST(
  alpha = 0.05,
  power = 0.80,
  delta = 0,
  sd = EQ2_HR$effsize$SE[[1]]*sqrt(EQ2_HR$smd$ntilde),
  eqb = 5,
  type = "paired"
)

power_EQ1_SP <- power_t_TOST(
  alpha = 0.05,
  power = 0.80,
  delta = 0,
  sd = EQ1_SP$effsize$SE[[1]]*sqrt(EQ1_SP$smd$ntilde),
  eqb = 0.5,
  type = "paired"
)

power_EQ2_SP <- power_t_TOST(
  alpha = 0.05,
  power = 0.80,
  delta = 0,
  sd = EQ2_SP$effsize$SE[[1]]*sqrt(EQ2_SP$smd$ntilde),
  eqb = 0.5,
  type = "paired"
)

# sample size calculation for population difference of 2 bpm or 0.2 km/h
power_EQ1_HR_2 <- power_t_TOST(
  alpha = 0.05,
  power = 0.80,
  delta = 2,
  sd = EQ1_HR$effsize$SE[[1]]*sqrt(EQ1_HR$smd$ntilde),
  eqb = 5,
  type = "paired"
)

power_EQ2_HR_2 <- power_t_TOST(
  alpha = 0.05,
  power = 0.80,
  delta = 2,
  sd = EQ2_HR$effsize$SE[[1]]*sqrt(EQ2_HR$smd$ntilde),
  eqb = 5,
  type = "paired"
)

power_EQ1_SP_2 <- power_t_TOST(
  alpha = 0.05,
  power = 0.80,
  delta = 0.2,
  sd = EQ1_SP$effsize$SE[[1]]*sqrt(EQ1_SP$smd$ntilde),
  eqb = 0.5,
  type = "paired"
)

power_EQ2_SP_2 <- power_t_TOST(
  alpha = 0.05,
  power = 0.80,
  delta = 0.2,
  sd = EQ2_SP$effsize$SE[[1]]*sqrt(EQ2_SP$smd$ntilde),
  eqb = 0.5,
  type = "paired"
)


# create an overview table
# this is a bit of a mess, but it works
samplsize_overview <- data.frame(
  c("BP1-GET HR","BP2-RCP HR", "BP1-GET Speed", "BP2-RCP Speed"),
  c(
    round(EQ1_HR$effsize$SE[[1]]*sqrt(EQ1_HR$smd$ntilde),2),
    round(EQ2_HR$effsize$SE[[1]]*sqrt(EQ2_HR$smd$ntilde),2),
    round(EQ1_SP$effsize$SE[[1]]*sqrt(EQ1_SP$smd$ntilde),2),
    round(EQ2_SP$effsize$SE[[1]]*sqrt(EQ2_SP$smd$ntilde),2)),
  c(
    round(power_EQ1_HR$n,0),
    round(power_EQ2_HR$n,0),
    round(power_EQ1_SP$n,0),
    round(power_EQ2_SP$n,0)
  ),
  c(
    round(power_EQ1_HR_2$n,0),
    round(power_EQ2_HR_2$n,0),
    round(power_EQ1_SP_2$n,0),
    round(power_EQ2_SP_2$n,0)
  )
)

samplsize_overview %>%
  kbl(
    col.names = c("Comparison", "Sample SD of the difference", "n for 0 difference", "n for small difference")) %>%
  kable_styling(full_width = FALSE)
```

Note: The first sample size calculation assumes a true difference between methods of 0. The second calculation assumes a true mean difference of 2 bpm for heart rate or 0.2 km/h for speed. Also, we assumed that our sample standard deviation of the difference approximates the population SD of the difference. This would only be the case for the specific estimation method used in our study, and, due to our sample size of n=11, is a low-precision estimate.

Based on these sample size calculations, we argue that future studies need at least 71 participants to test equivalence of BP2 with RCP an at least 128 participants to test equivalence of BP1 with GET to achieve 80% power *if the new method has the same standard deviation of differences*. If a future method for breakpoint estimation shows less variability (lower standard deviation of differences), lower sample sizes may suffice.

However, **we recommend future studies to specify their expected true difference in advance based on the existing literature.** It is unlikely that the differences is 0, thus resorting to the estimated sample sizes for a true difference of 0 may be perceived as bad practice. As seen in the above table, the required sample sizes are much lower when assuming a true difference of 0, compared to a small difference. This will lead to under-powered studies not able to severely test the hypothesis of equivalence between threshold detection methods.

## Summary

The four equivalence tests are summarized in the following table:

```{r}
#| label: TblSummary

options(scipen = 4)

# create the summary data frame
# not the cleanest code but rather manually defining
# all values
summary <- data.frame(
  a = c("BP1-GET HR (bpm)","BP1-GET Speed (km/h)", "BP2-RCP HR (bpm)", "BP2-RCP Speed (km/h)"), # comparison
  b = c(EQ1_HR$TOST[1,4], EQ1_SP$TOST[1,4], EQ2_HR$TOST[1,4],EQ2_SP$TOST[1,4]), # NHST
  c = c(EQ1_HR$TOST[2,4], EQ1_SP$TOST[2,4], EQ2_HR$TOST[2,4],EQ2_SP$TOST[2,4]), # Equalence test
  d = c(EQ1_HR$effsize[1,1], EQ1_SP$effsize[1,1], EQ2_HR$effsize[1,1],EQ2_SP$effsize[1,1]), # effect size
  e = c(EQ1_HR$effsize[1,3], EQ1_SP$effsize[1,3], EQ2_HR$effsize[1,3],EQ2_SP$effsize[1,3]), # CI_lower
  f = c(EQ1_HR$effsize[1,4], EQ1_SP$effsize[1,4], EQ2_HR$effsize[1,4],EQ2_SP$effsize[1,4]) # CI_upper
)

# and print as a kable
summary %>%
  mutate(
    across(c(b,c), ~ round(.x,4)),
    across(c(d,e,f), ~ round(.x,2))
  ) %>%
  kbl(col.names = c(
    "Comparison", "p NHST", "p equivalence test", "difference", "lower CI", "upper CI"
  )) %>%
  kable_styling()
```

Based on our previous stated hypotheses, we conclude:

**Hypothesis BP1-GET**

Both heart rate (p = `r round(EQ1_HR$TOST[2,4],3)`) and speed (p = `r round(EQ1_SP$TOST[2,4],3)`) at the first threshold are not equivalent. Further, the null hypotheses test shows that heart rate and speed are statistically different between BP1 and GET (p = `r round(EQ1_SP$TOST[1,4],3)` and p = `r round(EQ1_HR$TOST[1,4],3)`, respectively).

The SmO~2~ BP1 is not only not equivalent to GET, but also statistically different.

**Hypothesis BP2-RCP**

Similarly, both heart rate (p = `r round(EQ2_HR$TOST[2,4],3)`) and speed (p = `r round(EQ2_SP$TOST[2,4],3)`) at the second threshold are not equivalent. Further, the null hypotheses test shows that heart rate and speed are statistically different between BP1 and GET (p = `r round(EQ2_SP$TOST[1,4],4)` and p = `r round(EQ2_HR$TOST[1,4],4)`, respectively).

The SmO~2~ BP2 is not only not equivalent to RCP, but also statistically different.

This is visualized by plotting the mean differences and their 90% CIs (with or without individual data):

```{r}
#| label: FigSummarySimple
 
# combine some data
summary_data <- rbind(
  EQ1_HR$eff[1,],
  EQ1_SP$eff[1,],
  EQ2_HR$eff[1,],
  EQ2_SP$eff[1,]
)

# data for vlines
vlines <- data.frame(
  level = c("heart rate","heart rate","speed","speed"),
  vline_pos = c(-5, 5, -0.5, 0.5)
)

# create backgroud shading
shaded_df <- data.frame(
  level = c("heart rate", "speed"),
  xmin = c(-5,-0.5),
  xmax = c(5,0.5),
  ymin = -Inf,
  ymax = Inf
)

# x axis limits
xlimits <- data.frame(
  level = c("heart rate", "speed"),
  x_min = c(-16, -3),
  x_max = c(16, 3)
) %>%
  pivot_longer(cols = c(x_min, x_max),
               names_to = "limit_type",
               values_to = "x") 

# create plots for standard mean with 95% CI presentation
summary_data %>%
  mutate(
    level = c("heart rate","speed","heart rate","speed"),
    test = c("BP1-GET","BP1-GET", "BP2-RCP", "BP2-RCP")
  ) %>%
  ggplot(aes(estimate, test)) +
  geom_vline(data = vlines, aes(xintercept = vline_pos), linetype = 2) +
  geom_rect(data= shaded_df, aes (xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax), fill = "grey90", inherit.aes = FALSE) +
  geom_vline(xintercept = 0, linetype = 3) +
  geom_blank(data = xlimits, aes(x = x), inherit.aes = FALSE) +
  geom_point() +
  geom_linerange(aes(xmin = lower.ci, xmax = upper.ci)) +
  facet_wrap(~level, scales = "free") +
  labs(
    x = NULL,
    y = NULL,
    caption = "grey areas present the equivalence range"
  )
```

```{r}
#| label: FigSummaryIndividual

# data for vlines
vlines <- data.frame(
  Measure = c("heart rate","heart rate","speed","speed"),
  vline_pos = c(-5, 5, -0.5, 0.5)
)

# create backgroud shading
shaded_df <- data.frame(
  Measure = c("heart rate", "speed"),
  xmin = c(-5,-0.5),
  xmax = c(5,0.5),
  ymin = -Inf,
  ymax = Inf
)

# x axis limits
xlimits <- data.frame(
  Measure = c("heart rate", "speed"),
  x_min = c(-16, -3),
  x_max = c(16, 3)
) %>%
  pivot_longer(cols = c(x_min, x_max),
               names_to = "limit_type",
               values_to = "x") 

# create plots with density and individual data points
Study_data %>%
  mutate(diffHR1 = GET_HR - BP1_HR,
         diffSP1 = GET_speed - BP1_speed,
         diffHR2 = RCP_HR - BP2_HR,
         diffSP2 = RCP_speed - BP2_speed) %>%
  pivot_longer(
    cols = starts_with("diff"),
    names_to = "Comparison",
    values_to = "Difference",
    names_prefix = "diff"
  ) %>%
  mutate(Measure = ifelse(str_starts(Comparison, "SP"),"speed", "heart rate"),
         Threshold = ifelse(str_ends(Comparison, "1"),"1st Thd","2nd Thd"),
         Threshold = factor (Threshold, levels = c("2nd Thd","1st Thd"))) %>%
  ggplot(aes(Difference, Threshold, fill = Threshold)) +
  stat_slab(normalize = "panels", adjust = 1.5) +
  geom_rect(data= shaded_df, 
            aes (xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax), 
            fill = "grey90", alpha = 0.5, inherit.aes = FALSE) +
  stat_dots(color = clrs[3], fill = clrs[3], scale = 0.5) +
  geom_vline(data = vlines, aes(xintercept = vline_pos), linetype = 2) +
  geom_vline(xintercept = 0, linetype = 3) +
  stat_summary(fun = mean, geom = "point", color = "black", size = 2, position = position_nudge(y = -0.03)) +
  stat_summary(fun.data = mean_cl_normal, fun.args = list(conf.int = 0.90),
               geom = "pointrange", color = "black", position = position_nudge(y = -0.03)) +
  facet_wrap(~Measure, scales = "free") +
  labs(
    y = NULL,
    caption = "grey areas present the equivalence range"
  ) + 
  scale_color_manual(values = c(clrs[1:2])) +
  scale_fill_manual(values = c(clrs[1:2])) +
  theme(
    legend.position = "none"
    )
```

## Content for the manuscript {#sec-manuscript}

### Figure 1

```{r}
#| label: Fig1
#| output: false

#load the raw data for participants ID_03 and ID_06
ID3_data <- read_xlsx("data/input data/ID_03.xlsx")
ID6_data <- read_xlsx("data/input data/ID_06.xlsx")

pID3 <- ID3_data %>%
  ggplot(aes(Time,SmO2_right)) +
  geom_point(size = 2) +
  # now add vertical lines for the BPs and annotate them
  geom_vline(xintercept = as.numeric(Study_data %>% filter(ID == 3) %>% select(BP1_time,BP2_time)),
             linetype = 2, linewidth = 1,
             color = clrs[1]) +
  annotate("text",
           x = as.numeric(Study_data %>% filter(ID == 3) %>% select(BP1_time,BP2_time))+10,
           y = 100,
           label = c("BP1","BP2"),
           hjust = 0,
           color = clrs[1]) +
  # repeat for VTs
  geom_vline(xintercept = as.numeric(Study_data %>% filter(ID == 3) %>% select(GET_time,RCP_time)),
             linetype = 2, linewidth = 1,
             color = clrs[2]) +
  annotate("text",
           x = as.numeric(Study_data %>% filter(ID == 3) %>% select(GET_time,RCP_time))+10,
           y = 95,
           label = c("GET","RCP"),
           hjust = 0,
           color = clrs[2]) +
  scale_y_continuous(limits = c(0,100), breaks = seq(0,100,20)) +
  scale_x_continuous(limits = c(0,1500), breaks = seq(0,1470,240)) +
  labs(
    x = "Time [s]",
    y = expression(bold(SmO[2]~"[%]")),
    caption = "ID03"
  ) +
  theme(
    axis.title = element_text(family = "Arial", size = 9),
    axis.text = element_text(family = "Arial", size = 9),
    text = element_text(family = "Arial", size = 9)
  )

pID6 <- ID6_data %>%
  ggplot(aes(Time,SmO2_right)) +
  geom_point(size = 2) +
  # now add vertical lines for the BPs and annotate them
  geom_vline(xintercept = as.numeric(Study_data %>% filter(ID == 6) %>% select(BP1_time,BP2_time)),
             linetype = 2, linewidth = 1,
             color = clrs[1]) +
  annotate("text",
           x = as.numeric(Study_data %>% filter(ID == 6) %>% select(BP1_time,BP2_time))+10,
           y = 100,
           label = c("BP1","BP2"),
           hjust = 0,
           color = clrs[1]) +
  # repeat for VTs
  geom_vline(xintercept = as.numeric(Study_data %>% filter(ID == 6) %>% select(GET_time,RCP_time)),
             linetype = 2, linewidth = 1,
             color = clrs[2]) +
  annotate("text",
           x = as.numeric(Study_data %>% filter(ID == 6) %>% select(GET_time,RCP_time))+10,
           y = 95,
           label = c("GET","RCP"),
           hjust = 0,
           color = clrs[2]) +
  scale_y_continuous(limits = c(0,100), breaks = seq(0,100,20)) +
  scale_x_continuous(limits = c(0,1500), breaks = seq(0,1260,240)) +
  labs(
    x = "Time [s]",
    y = expression(bold(SmO[2]~"[%]")),
    caption = "ID06"
  ) +
  theme(
    axis.title = element_text(family = "Arial", size = 9),
    axis.text = element_text(family = "Arial", size = 9),
    text = element_text(family = "Arial", size = 9)
  )

# now combine plots next to each other
Fig1 <- pID3 + pID6 + 
  patchwork::plot_annotation(tag_levels = "A")

ggsave("output/results/Fig1.jpeg",
       plot = Fig1,
       device = "jpeg",
       width = 300,
       height = 112.5,
       units = "mm",
       dpi = 1200)

ggsave("output/results/Fig1.eps",
       plot = Fig1,
       device = cairo_ps,
       width = 300,
       height = 112.5,
       units = "mm",
       dpi = 1200)
```

### Figure 2

```{r}
#| label: Fig2

# data for vlines
vlines <- data.frame(
  Measure = c("Heart Rate [bpm]","Heart Rate [bpm]","Speed [km/h]","Speed [km/h]"),
  vline_pos = c(-5, 5, -0.5, 0.5)
)

# create backgroud shading
shaded_df <- data.frame(
  Measure = c("Heart Rate [bpm]", "Speed [km/h]"),
  xmin = c(-5,-0.5),
  xmax = c(5,0.5),
  ymin = -Inf,
  ymax = Inf
)

# x axis limits
xlimits <- data.frame(
  Measure = c("Heart Rate [bpm]", "Speed [km/h]"),
  x_min = c(-16, -3),
  x_max = c(16, 3)
) %>%
  pivot_longer(cols = c(x_min, x_max),
               names_to = "limit_type",
               values_to = "x") 


Fig2 <- Study_data %>%
  mutate(diffHR1 = GET_HR - BP1_HR,
         diffSP1 = GET_speed - BP1_speed,
         diffHR2 = RCP_HR - BP2_HR,
         diffSP2 = RCP_speed - BP2_speed) %>%
  pivot_longer(
    cols = starts_with("diff"),
    names_to = "Comparison",
    values_to = "Difference",
    names_prefix = "diff"
  ) %>%
  mutate(Measure = ifelse(str_starts(Comparison, "SP"),"Speed [km/h]", "Heart Rate [bpm]"),
         Threshold = ifelse(str_ends(Comparison, "1"),"1st Thd","2nd Thd"),
         Threshold = factor (Threshold, levels = c("2nd Thd","1st Thd"))) %>%
  ggplot(aes(Difference, Threshold, fill = Threshold)) +
  stat_slab(normalize = "panels", adjust = 1.5) +
  geom_rect(data= shaded_df, 
            aes (xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax), 
            fill = "grey90", alpha = 0.5, inherit.aes = FALSE) +
  stat_dots(color = "grey30", fill = "grey30", scale = 0.5) +
  geom_vline(data = vlines, aes(xintercept = vline_pos), linetype = 2) +
  geom_vline(xintercept = 0, linetype = 3) +
  stat_summary(fun = mean, geom = "point", color = "black", size = 2, position = position_nudge(y = -0.03)) +
  stat_summary(fun.data = mean_cl_normal, fun.args = list(conf.int = 0.90),
               geom = "pointrange", color = "black", position = position_nudge(y = -0.03)) +
  facet_wrap(~Measure, scales = "free") +
  labs(
    y = NULL,
    x = "Difference (VT minus BP)"
  ) + 
  scale_color_manual(values = c(clrs[1:2])) +
  scale_fill_manual(values = c(clrs[1:2])) +
  theme(
    legend.position = "none",
    axis.title = element_text(family = "Arial", size = 9),
    axis.text = element_text(family = "Arial", size = 9),
    text = element_text(family = "Arial", size = 9)
    )

ggsave("output/results/Fig2.jpeg",
       plot = Fig2,
       device = "jpeg",
       width = 150,
       height = 112.5,
       units = "mm",
       dpi = 1200)

ggsave("output/results/Fig2.eps",
       plot = Fig1,
       device = cairo_ps,
       width = 150,
       height = 112.5,
       units = "mm",
       dpi = 1200)
```

### Table 1

```{r}
#| label: Tbl1
#| output: false

Tbl1 <- Study_data %>%
  select(Age, Height, Weight, ATT_left, ATT_right, v_peak, MAS, absVO2max, VO2max) %>%
  mutate(BMI = Weight / (Height/100)^2) %>%
    tbl_summary(
    type = ~ "continuous",
    statistic = all_continuous() ~ "{mean} \u00B1 {sd} [{min} - {max}]",
    missing_text = "missing",
    digits = list(
      c(Age, Height, Weight, ATT_left, ATT_right, v_peak, MAS, VO2max, BMI) ~ 1,
      c(absVO2max) ~ 0
    )
  ) %>%
  as_gt() %>%
  tab_options(table.font.size = px(12)) %>%
  opt_table_font(font = list(gt::google_font("Arial")))

# and export in various formats

# html
gtsave(Tbl1, "output/results/Tbl1.html")

# html
gtsave(Tbl1, "output/results/Tbl1.png")

# latex
gtsave(Tbl1, "output/results/Tbl1.tex")

# word
gtsave(Tbl1, "output/results/Tbl1.docx")
```

## Session Info

```{r}
sessionInfo()
```

## References
